Peer Assessment 1
=================

The following document is my answer for *Peer Assessment 1* during the course *Reproducible Research*. 

## Loading and preprocessing the data

As a first step let us load a dataset dedicated for this assessment and look for some details of it.

```{r}
activity <- read.csv("/Volumes/Obraz dysku/2015.02 Coursera/AS1/activity.csv")
activity$dateD <- as.Date(as.character(activity$date), "%Y-%m-%d")
summary(activity)
str(activity)
```

## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day

I can use the fact that now in my dataset dates are stored as factors, so the total 
number of steps taken per each day is:
```{r}
StepsPerDay <- sapply(X = split(x = activity$steps, f = activity$date), 
                      FUN = function(x) sum(x, na.rm = TRUE))
print(StepsPerDay)
```

### A histogram of the total number of steps taken each day

```{r}
hist(StepsPerDay)
```

### The mean and median of the total number of steps taken per day

The mean value of the total number of steps taken per day is:
```{r}
mean(StepsPerDay)
```

The median value of the total number of steps taken per day is:
```{r}
median(StepsPerDay)
```

## What is the average daily activity pattern?

A time series plot of the 5-minute interval (x-axis) and the average number of 
steps taken, averaged across all days (y-axis)

```{r}
x <- sapply(X = split(x = activity$steps, f = as.factor(activity$interval)), 
            FUN = function(x) mean(x, na.rm = TRUE))
plot(x = as.numeric(names(x)), y = x, type = "l", xlab = "5-minute interval", ylab = "av. number of steps")
```

The 5-minute interval, which on average across all the days in the dataset, 
contains the maximum number of steps has a number

```{r}
names(x)[x==max(x)]
```


## Imputing missing values

The total number of missing values in the dataset is
```{r}
sum(!complete.cases(activity))
```

Imputing missing values I will use a mean value of steps for the 5-minute interval over all days.

```{r}
activityImp <- activity
activityImp$steps[is.na(activityImp$steps)] <- x[as.character(activityImp$interval[is.na(activityImp$steps)])]
summary(activityImp$steps)
```

We make now a histogram of the total number of steps taken each day and we calculate and report the mean and median total number of steps taken per day:
```{r}
StepsPerDayImp <- sapply(X = split(x = activityImp$steps, f = activityImp$date), 
                      FUN = function(x) sum(x, na.rm = TRUE))
mean(StepsPerDayImp)
median(StepsPerDayImp)
hist(StepsPerDayImp)
```

## Are there differences in activity patterns between weekdays and weekends?

First we create a new factor variable in the dataset with two levels: **weekday** and 
**weekend**, indicating whether a given date is a weekday or weekend day (*I used a 
polish version of R; "sobota" = Saturday, "niedziela" = Sunday*).

```{r}
activity$dayF <- weekdays(activity$dateD)
activity$dayF[activity$dayF=="sobota" | activity$dayF=="niedziela"] <- "weekend"
activity$dayF[activity$dayF!="weekend"] <- "weekday"
summary(as.factor(activity$dayF))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute 
interval (x-axis) and the average number of steps taken, averaged across all weekday 
days or weekend days (y-axis). See the README file in the GitHub repository to see 
an example of what this plot should look like using simulated data.

```{r}
activityWD <- activity[activity$dayF == "weekday",]
xWD <- sapply(X = split(x = activityWD$steps, f = as.factor(activityWD$interval)), 
            FUN = function(x) mean(x, na.rm = TRUE))

activityWE <- activity[activity$dayF == "weekend",]
xWE <- sapply(X = split(x = activityWE$steps, f = as.factor(activityWE$interval)), 
            FUN = function(x) mean(x, na.rm = TRUE))
par(mfcol = c(1, 2))
plot(x = as.numeric(names(xWD)), y = xWD, type = "l", main = "Weekdays",
     xlab = "5-minute interval", ylab = "av. number of steps")
plot(x = as.numeric(names(xWE)), y = xWE, type = "l", main = "Weekends", 
     xlab = "5-minute interval", ylab = "av. number of steps")
```




