Peer Assessment 1
=================

The following document is my answer for *Peer Assessment 1* during the course *Reproducible Research*. 

## Loading and preprocessing the data

As a first step let us load a dataset dedicated for this assessment and look for some details of it.

```{r}
activity <- read.csv("/Volumes/Obraz dysku/2015.02 Coursera/AS1/activity.csv")
activity$dateD <- as.Date(as.character(activity$date), "%Y-%m-%d")
summary(activity)
str(activity)
```

## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day

I can use the fact that now in my dataset dates are stored as factors, so the total 
number of steps taken per each day is:
```{r}
StepsPerDay <- sapply(X = split(x = activity$steps, f = activity$date), 
                      FUN = function(x) sum(x, na.rm = TRUE))
print(StepsPerDay)
```

### A histogram of the total number of steps taken each day

```{r}
hist(StepsPerDay)
```

### The mean and median of the total number of steps taken per day

The mean value of the total number of steps taken per day is:
```{r}
mean(StepsPerDay)
```

The median value of the total number of steps taken per day is:
```{r}
median(StepsPerDay)
```

## What is the average daily activity pattern?

A time series plot of the 5-minute interval (x-axis) and the average number of 
steps taken, averaged across all days (y-axis)

```{r}
x <- sapply(X = split(x = activity$steps, f = as.factor(activity$interval)), 
            FUN = function(x) mean(x, na.rm = TRUE))
plot(x = x, type = "l")
```

The 5-minute interval, which on average across all the days in the dataset, 
contains the maximum number of steps has a number

```{r}
names(x)[x==max(x)]
```


## Imputing missing values

The total number of missing values in the dataset is
```{r}
sum(!complete.cases(activity))
```

Imputing missing values I will use a mean value of steps for the 5-minute interval over all days.




## Are there differences in activity patterns between weekdays and weekends?

